---
title: "Amazon Jobs"
output: html_notebook
---

Let's check out the data!

```{r}
library(readr)
amazon <- read_csv("amazon_jobs_dataset.csv")
head(amazon)
```

Questions that we can focus on:
  - What are the popular locations of job openings?
  - Are there certain months when job opening are more? 
    - How many jobs are posted every year?
    - How many jobs are posted every month?
  - What are most popular job titles?
  - What are minimum educational qualifications?
    - Which job requires higher degree?
  - What languages are required for jobs?
  - How many years of experience is required?
    - Which job categories require more experience?
  - Word clouds of preferred qualifications
  - Word clouds of Basic Qualifications
  - Word clouds of JOb description

```{r}
dim(amazon)
```
```{r}
unique(amazon$Posting_date)[1:20]
```

We see that in some dates, there are multiple spaces between month and day. Let's split it to year month and day.

```{r}
library(stringr)
amazon$posting_month <- str_extract(amazon$Posting_date , "([A-Za-z]+)")
amazon$posting_day <- str_extract(amazon$Posting_date , "([0-9]+)")
amazon$posting_year <- str_extract(amazon$Posting_date, "([0-9]{4})")
```

```{r}
unique(amazon$posting_month)
unique(amazon$posting_day)
unique(amazon$posting_year)
```
Let's also extract the countries from location column.

```{r}
amazon$country <- str_extract(amazon$location, "[A-Za-z]+")
unique(amazon$country)
```

```{r}
unique(amazon$location)
```

```{r}
amazon$city <- word(amazon$location, -1)
unique(amazon$city)
```

Let us correct the city names that have multiple words.
```{r}
amazon$city <- gsub(pattern = "Alto", replacement = "East Palo Alto", amazon$city)
amazon$city <- gsub(pattern = "Ia<U+015F>i", replacement = "Iasi", amazon$city)
amazon$city <- gsub(pattern = "Monica", replacement = "Santa Monica", amazon$city)
amazon$city <- gsub(pattern = "Francisco", replacement = "San Francisco", amazon$city)
amazon$city <- gsub(pattern = "Wroclawskie", replacement = "Bielany Wroclawskie", amazon$city)
amazon$city <- gsub(pattern = "Diego", replacement = "San Diego", amazon$city)
amazon$city <- gsub(pattern = "Barbara", replacement = "Santa Barbara", amazon$city)
amazon$city <- gsub(pattern = "York", replacement = "New York", amazon$city)
amazon$city <- gsub(pattern = "Town", replacement = "Cape Town", amazon$city)
amazon$city <- gsub(pattern = "Cruz", replacement = "Santa Cruz", amazon$city)
amazon$city <- gsub(pattern = "Obispo", replacement = "San Luis Obispo", amazon$city)
amazon$city <- gsub(pattern = "Paulo", replacement = "Sao Paulo", amazon$city)
amazon$city <- gsub(pattern = "Hasharon", replacement = "Hod Hasharon", amazon$city)
amazon$city <- gsub(pattern = "IL,", replacement = "Hod Hasharon", amazon$city)
amazon$city <- gsub(pattern = "On,", replacement = "Ottawa", amazon$city)
amazon$city <- gsub(pattern = "Jose", replacement = "San Jose", amazon$city)
amazon$city <- gsub(pattern = "CA,", replacement = "Ottawa", amazon$city)
amazon$city <- gsub(pattern = "Vegas", replacement = "Las Vegas", amazon$city)
```

The word "City" has been used with three different cities, so include them.

```{r}
amazon$city <- ifelse(amazon$location == "US, CA, Foster City", "Foster City", amazon$city)
amazon$city <- ifelse(amazon$location == "MX, EMEX, Mexico City", "Mexico City", amazon$city)
amazon$city <- ifelse(amazon$location == "US, NJ, Jersey City", "Jersey City", amazon$city)
```

```{r}
head(amazon)
```

# What are most popular locations of Job Openings?

```{r}
library(dplyr)
location <- amazon[, c("country", "city")]
loc1 <- as.data.frame(table(location$city))
colnames(loc1) <- c("city", "Freq")

loc1 <- merge(loc1,location,by="city")
loc1 <- arrange(loc1, desc(Freq))
loc1 <- loc1 %>% distinct(city, .keep_all = TRUE)
loc1
```


```{r fig.width=5, fig.height=3, echo=FALSE}
library(ggplot2)
library(Rmisc)
tilt_theme <- theme(axis.text.x=element_text(angle=45, hjust=1))
p1 <- ggplot(data = head(loc1, 10), aes(city, Freq, fill = country)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Top 10 Job Locations") + 
      coord_flip() +
      tilt_theme 
p1
```

```{r fig.width=5, fig.height=3, echo=FALSE}
p2 <- ggplot(data = tail(loc1, 50), aes(city, Freq, fill = country)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Last 50 Job Locations") + 
      coord_flip() +
      tilt_theme 
p2
```

Distribution of Jobs in all countries except US
```{r}
p3 <- ggplot(filter(loc1,loc1$country!="US"), aes(x=factor(country), y=Freq, fill = factor(country))) + 
  geom_boxplot() 
p3
```

Distribution of jobs in US

```{r}
p4 <- ggplot(data = filter(loc1, loc1$country == "US"), aes(city, Freq, fill = city)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Jobs in US") + 
      coord_flip() +
      guides(fill=FALSE) +
      tilt_theme 
p4
```

# Analysis of Job Posting Date

## How many jobs are posted every year and every month? Is there a certain period of year when jobs are posted more fequently?

```{r}

amazon$Posting_date <- paste(amazon$posting_year, amazon$posting_month, amazon$posting_day,sep="-")

amazon$Posting_date <- as.Date(strptime(amazon$Posting_date,format="%Y-%b-%d"))
#amazon$Posting_date$zone <- NULL
(amazon$Posting_date[1:10])

#get day of the week based on date
dow <- function(x) format(as.Date(x), "%A")
amazon$weekday <- dow(amazon$Posting_date)
amazon$weekday[1:20]
```

```{r}
posting_date <- amazon[,c("posting_month", "posting_year", "weekday", "Posting_date")]
head(posting_date)
```

```{r}
year <- as.data.frame(table(posting_date$posting_year))
colnames(year) <- c("Year", "Freq")
p5 <- ggplot(data = year, aes(Year, Freq, fill = Year)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Jobs postings by Year") + 
      guides(fill=FALSE) +
      tilt_theme 
p5
```

```{r}
mon <- as.data.frame(table(posting_date$posting_month))
colnames(mon) <- c("Mon", "Freq")
lev = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
mon$Month <- factor(mon$Mon, lev)
 
p6 <- ggplot(data = mon, aes(Month, Freq, fill = Month)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Jobs postings by Month") + 
      guides(fill=FALSE) +
      tilt_theme 
p6
```

```{r}
jday <- as.data.frame(table(posting_date$weekday))
colnames(jday) <- c("wd", "Freq")
lev <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
jday$weekday <- factor(jday$wd, lev)

p7 <- ggplot(data = jday, aes(weekday, Freq, fill = weekday)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Jobs postings by Weekday") + 
      guides(fill=FALSE) +
      tilt_theme 
p7
```

The above graphs show that most of the job postings are recent and there has been large increase in jobs in 2017 and 2018. January and february are hot months for jobs. Also, the jobs are posted generally on weekdays, mostly on tuesdays and wednesday. This is true as generally employees are completing more work in the first half of weekdays. 

# Analysis of Job titles

## What are most popular job titles?

```{r}
title <- arrange(as.data.frame(table(amazon$Title)), desc(Freq))
colnames(title) <- c("Title", "Freq")
p8 <- ggplot(data = head(title, 10), aes(Title, Freq, fill = Title)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Top 10 jobs") + 
      guides(fill=FALSE) +
      coord_flip() +
      tilt_theme 
p8
```

There are various job postings. Top 10 are shown above. Let us now divide it in different domains and different positions.
```{r}

# software, game, quality, data, web, security, sale,iot, research
# lead, manager, developer, engineer, consultant, artist, analyst, scientist,
domain <- c("software", "game", "quality", "data", "web", "security", "sale", "research", "iot", "ui")
positions <- c("lead", "leader","manager", "developer", "engineer", "consultant", "artist", "analyst", "scientist")
amazon$Title <- tolower(amazon$Title)
#unique(amazon$Title)
amazon$domain <- NA
amazon$positions <- NA

```



```{r}
for(i in domain){
amazon$domain1 <- grepl(i, amazon$Title)
amazon$domain <- ifelse(amazon$domain1 == "TRUE", i, amazon$domain)
}
amazon$domain <- ifelse(amazon$domain == FALSE, "other", amazon$domain)

for(i in positions){
  amazon$pos1 <- grepl(i, amazon$Title)
  amazon$positions <- ifelse(amazon$pos1 == "TRUE", i, amazon$positions)
}
amazon$positions <- ifelse(is.na(amazon$positions), "other", amazon$positions)
amazon <- subset(amazon, select = -c(domain1, pos1))

```

```{r}
pos <- as.data.frame(table(amazon$positions))
colnames(pos) <- c("position", "Freq")
p9 <- ggplot(data = pos, aes(position, Freq, fill = position)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Top 10 domains") + 
      guides(fill=FALSE) +
      coord_flip() 
p9
```

```{r}
dom <- as.data.frame(table(amazon$domain))
colnames(dom) <- c("domain", "Freq")
p9 <- ggplot(data = dom, aes(domain, Freq, fill = domain)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Top 10 domains") + 
      guides(fill=FALSE) +
      coord_flip() 
p9
```

# What are minimum educational qualifications?

```{r}
amazon$education <- NA
degree_list = c("ba", "bs", "b.tech","bachelor", "phd","ms","master", "mba","m.tech")
amazon$`BASIC QUALIFICATIONS` <- tolower(amazon$`BASIC QUALIFICATIONS`)

for(i in degree_list){
  amazon$deg1 <- grepl(i, amazon$`BASIC QUALIFICATIONS`)
  amazon$education <- ifelse(amazon$deg1 == "TRUE", i, amazon$education)
}
amazon$education <- ifelse(is.na(amazon$education), "other", amazon$education)
amazon <- subset(amazon, select = -c(deg1))

```

```{r}
edu <- as.data.frame(table(amazon$education))
colnames(edu) <- c("Education", "Freq")
p10 <- ggplot(data = edu, aes(Education, Freq, fill = Education)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Jobs based on degrees") + 
      guides(fill=FALSE) +
      coord_flip() 
p10
```

Basic Qualifications generally requires Masters degree.

## Which jobs require which type of degrees?


```{r fig.width=7, fig.height=5, echo=FALSE}
counts <- as.data.frame(table(amazon$domain, amazon$education))
colnames(counts) <- c("Domain", "Education", "Freq")
p11 <- ggplot(data = filter(counts, counts$Domain != "software"), aes(Domain, Freq)) +
      geom_bar(aes(fill = Education), position = "dodge",stat = "identity") +
      ggtitle("Domain distribution based on Education") 

p12 <- ggplot(data = filter(counts, counts$Domain == "software"), aes(Domain, Freq)) +
      geom_bar(aes(fill = Education), position = "dodge",stat = "identity") +
      ggtitle("Domain distribution based on Education") 
multiplot(p11, p12)
```

```{r fig.width=7, fig.height=5, echo=FALSE}
counts <- as.data.frame(table(amazon$positions, amazon$education))
colnames(counts) <- c("Positions", "Education", "Freq")
p13 <- ggplot(data = filter(counts, counts$Positions != "engineer"), aes(Positions, Freq)) +
      geom_bar(aes(fill = Education), position = "dodge",stat = "identity") +
      ggtitle("Positions distribution based on Education") 

p14 <- ggplot(data = filter(counts, counts$Positions == "engineer"), aes(Positions, Freq)) +
      geom_bar(aes(fill = Education), position = "dodge",stat = "identity") +
      ggtitle("Positions distribution based on Education") 
multiplot(p13, p14)
```

# What languages are in demand?

```{r}
list_of_occurence <- c()
languages <- c('swift','matlab','mongodb','hadoop','cosmos', 'mysql','spark', 'pig', 'python', 'java.', 'java,','c[++]', 'php', 'javascript', 'objective c', 'ruby', 'perl','c ','c#', ' r,')
amazon$`PREFERRED QUALIFICATIONS` <- tolower(amazon$`PREFERRED QUALIFICATIONS`)

for(i in languages){
  amazon$dummy <- str_count(amazon$`PREFERRED QUALIFICATIONS`, i)
  list_of_occurence <- c(list_of_occurence, sum(amazon$dummy, na.rm = TRUE))
}
list_of_occurence
```

```{r}
lan <- data.frame(cbind(languages, as.numeric(list_of_occurence)))
p15 <- ggplot(data = lan, aes(languages, list_of_occurence, fill = languages)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=list_of_occurence), color="black", size=3) +
      ggtitle("Languages in job description") + 
      guides(fill=FALSE) +
      coord_flip() 
p15
```

Java, C, C++ and python are popular on-demand languages in jobs.


# How many years of experience is required?

```{r}
amazon$dummy <- str_extract(amazon$`BASIC QUALIFICATIONS`, '([0-9]+) year')
amazon$dummy <- ifelse(is.na(amazon$dummy), "0 year", amazon$dummy)

split_year <- function(var){
 return(as.numeric(unlist(str_split(var, " "))[1]))
}
amazon$experience <- sapply(amazon$dummy, split_year)

```

Let's see the experience, excluding 0 year experience.
```{r}
exp <- as.data.frame(table(amazon$experience))
colnames(exp) <- c("Experience", "Freq")
p16 <- ggplot(data = filter(exp, exp$Experience != 0), aes(Experience, Freq, fill = Experience)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Freq), color="black", size=3) +
      ggtitle("Experience needed") + 
      guides(fill=FALSE) +
      coord_flip() 
p16
```

```{r fig.width=7, fig.height=5, echo=FALSE}
counts <- as.data.frame(table(amazon$positions, amazon$experience))
colnames(counts) <- c("Positions", "Experience", "Freq")
p17 <- ggplot(data = filter(counts, counts$Experience != 0 & counts$Positions != "engineer"), aes(Positions, Freq)) +
      geom_bar(aes(fill = Experience), position = "dodge",stat = "identity") +
      ggtitle("Positions distribution based on Experience for positions excluding enginner") 
p18 <- ggplot(data = filter(counts, counts$Experience == 0 & counts$Positions != "engineer"), aes(Positions, Freq)) +
      geom_bar(aes(fill = Experience), position = "dodge",stat = "identity") +
      ggtitle("Positions where Experience is not defined excluding engineer") 
p19 <- ggplot(data = filter(counts, counts$Positions == "engineer"), aes(Positions, Freq)) +
      geom_bar(aes(fill = Experience), position = "dodge",stat = "identity") +
      ggtitle("Engineer Positions distribution based on Experience") 
p20 <- ggplot(data = filter(counts, counts$Experience != 0 & counts$Positions == "engineer"), aes(Positions, Freq)) +
      geom_bar(aes(fill = Experience), position = "dodge",stat = "identity") +
      ggtitle("Engineer Positions where Experience is not zero") 
multiplot(p17, p18, p19, p20, layout = matrix(c(1,2,3,4), nrow=2, byrow=TRUE))

```

```{r fig.width=7, fig.height=5, echo=FALSE}
counts <- as.data.frame(table(amazon$domain, amazon$experience))
colnames(counts) <- c("Domain", "Experience", "Freq")
p21 <- ggplot(data = filter(counts, counts$Experience != 0 & counts$Domain != "software"), aes(Domain, Freq)) +
      geom_bar(aes(fill = Experience), position = "dodge",stat = "identity") +
      ggtitle("Domain distribution based on Experience for domain excluding software") 

p22 <- ggplot(data = filter(counts, counts$Experience == 0 & counts$Domain != "software"), aes(Domain, Freq)) +
      geom_bar(aes(fill = Experience), position = "dodge",stat = "identity") +
      ggtitle("Domain where Experience is not defined for domain excluding software") 

p23 <- ggplot(data = filter(counts, counts$Domain == "software"), aes(Domain, Freq)) +
      geom_bar(aes(fill = Experience), position = "dodge",stat = "identity") +
      ggtitle("Domain distribution based on Experience for software") 

p24 <- ggplot(data = filter(counts, counts$Experience != 0 & counts$Domain == "software"), aes(Domain, Freq)) +
      geom_bar(aes(fill = Experience), position = "dodge",stat = "identity") +
      ggtitle("Domain where Experience is not zero for software") 

multiplot(p21, p22, p23, p24, layout = matrix(c(1,2,3,4), nrow=2, byrow=TRUE))
```

# Word Clouds of Qualifications and Job description

```{r}
library(wordcloud)
library(SnowballC)
library(RColorBrewer)
library(tm)
texts <- amazon$`PREFERRED QUALIFICATIONS`
#texts <- iconv(texts, to = "utf-8")
corpus <- Corpus(VectorSource(texts))
corpus <- tm_map(corpus, PlainTextDocument)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeWords, stopwords('english'))
#corpus <- tm_map(corpus, stemDocument)
corpus <- tm_map(corpus, removeWords, c("00b7","and", "this", "there")) 
corpus <- Corpus(VectorSource(corpus))
dtm <- TermDocumentMatrix(corpus)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
d <- d[-which(d$word %in% c("00b7","and","this","that")),]
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Preferred Qualifications Word Cloud")
```

```{r}
texts <- amazon$`BASIC QUALIFICATIONS`
#texts <- iconv(texts, to = "utf-8")
corpus <- Corpus(VectorSource(texts))
corpus <- tm_map(corpus, PlainTextDocument)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeWords, stopwords('english'))
#corpus <- tm_map(corpus, stemDocument)
corpus <- tm_map(corpus, removeWords, c("00b7","and", "this", "there")) 
corpus <- Corpus(VectorSource(corpus))
dtm <- TermDocumentMatrix(corpus)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
d <- d[-which(d$word %in% c("00b7","and","this","that")),]
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Basic Qualifications Word Cloud")
```

```{r  fig.width=5, fig.height=5, echo=FALSE}
texts <- amazon$DESCRIPTION
#texts <- iconv(texts, to = "utf-8")
corpus <- Corpus(VectorSource(texts))
corpus <- tm_map(corpus, PlainTextDocument)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeWords, stopwords('english'))
#corpus <- tm_map(corpus, stemDocument)
corpus <- tm_map(corpus, removeWords, c("00b7","and", "this", "there", "you", "will")) 
corpus <- Corpus(VectorSource(corpus))
dtm <- TermDocumentMatrix(corpus)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
d <- d[-which(d$word %in% c("00b7","and","this","that", "you", "will")),]
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Job Description Word Cloud")
```

```{r}
texts <- amazon$Title
#texts <- iconv(texts, to = "utf-8")
corpus <- Corpus(VectorSource(texts))
corpus <- tm_map(corpus, PlainTextDocument)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeWords, stopwords('english'))
#corpus <- tm_map(corpus, stemDocument)
corpus <- tm_map(corpus, removeWords, c("00b7","and", "this", "there", "you", "will")) 
corpus <- Corpus(VectorSource(corpus))
dtm <- TermDocumentMatrix(corpus)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
#d <- d[-which(d$word %in% c("00b7","and","this","that", "you", "will")),]
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Job Title Word Cloud")
```


